%*******************************************************
% Acknowledgments
%*******************************************************
%\pdfbookmark[1]{Acknowledgements}{acknowledgements}
\chapter*{Acknowledgements}

I would like to express my deep gratitude to all the people who supported me during this work.

First and foremost, I am grateful to my supervisor Dr Tobias Marschall who introduced me to this haplotyping problem.
He helped me to develop problem solving skills for solving a research problem
and provided me his able guidance in completing this thesis work. 
Thanks a million to Prof Volkhard Helms for his consent to give time to review this thesis at a short notice.
I would like to offer special thanks to Prof Thomas Lengauer for providing me his valuable remarks on an earlier version of some of the chapters of this thesis.
He provided me financial support and guided me about ways to pursue my passion.
I am thankful to Prof George Church for his valuable comments on an earlier version of this thesis.

I am deeply grateful to Prof Richard Durbin whose lab I visited for my research internship.
His enthusiasm for genomics and guidance on future career prospects largely motivated me.
Dr Tobias Mömke has been great source of inspiration and advice for me.
I am continually amazed by his problem solving ability, which provided me motivation to solve ambitious problems.

I am immensely thankful to all collaborators and coauthors for their support and assistance.
It was great working with David Porubsky and Ashley Sanders, and learning the magic of Strand-Seq technology.
Thanks to Mikko Rautiainen and vgteam for great discussions on the graph world of genomics.

I appreciate the discussions with my office-mate Ali Ghaffaari on violin and classical music.
Special thanks to Jana for always providing a friendly support and guidance.
I would like to extend my thanks to my group mates, who organized different events at weekends. The events were enjoyable experiences.

Jana, Fabio, Lara, Prabhav and Adam proofread a few chapters of this thesis and I greatly appreciate their valuable remarks and feedback.

I appreciate the travel financial support from Michelle Carnell, technical help from Achim, George and MPII-IST and other administrative help from Susanne and Ruth.  

Finally, I would like to express the gratitude to my family and friends, who greatly helped me in this journey.





%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End:
